`ro warning=FALSE, message=FALSE, comment=NA, tidy=FALSE, cache=TRUE, verbose=TRUE, cache.path="2012-05-15-nullfallacy/" or`

# Code for Prosecutors Fallacy 

Load the required libraries
 
``` {r libraries}
require(populationdynamics)
require(earlywarning)
require(reshape2)
require(snowfall)
````

Simulate a dataset from the full individual, nonlinear model, with stable parameters (*.e.g.* not approaching a bifurcation).

This defines our simulation function

``` {r simdatf_null}
select_crashes <- function(n){
	T<- 5000
	n_pts <- n
	pars = c(Xo = 500, e = 0.5, a = 180, K = 1000, h = 200,
    i = 0, Da = 0, Dt = 0, p = 2)
	sn <- saddle_node_ibm(pars, times=seq(0,T, length=n_pts), reps=500)
	d <- dim(sn$x1)
	crashed <- which(sn$x1[d[1],]==0)
#	sn$x1[,crashed] 
	sn$x1[1:501,]
}
````

To take advantage of parallelization, we loop over this function a set number of times.  

``` {r simdat_null}
sfInit(parallel=TRUE, cpu=12)
sfLibrary(populationdynamics)
sfExportAll()
examples <-  sfLapply(1:24, function(i) select_crashes(50000))
dat <- melt(as.matrix(as.data.frame(examples, check.names=FALSE)))
#examples <- select_crashes(50000)
dat <- melt(examples)
names(dat) = c("time", "reps", "value")
levels(dat$reps) <- 1:length(levels(dat$reps)) # use numbers for reps
````

Zoom in on the relevant area of data near the crash

``` {r subsetdata_null }
require(plyr)
nullzoom <- ddply(dat, "reps", function(X){
#    tip <- min(which(X$value==0))
#    index <- max(tip-500,1):tip
#    data.frame(time=X$time[index], value=X$value[index])
    data.frame(time=X$time, value=X$value)
    })
save(list="nullzoom", file="nullzoom.rda")
````

A plot of the first 9 datasets over the interval used for the warning signal calculation.

``` {r replicate_crashes_null}
require(ggplot2)
ggplot(nullzoom) + geom_line(aes(time, value, group=reps), alpha=.1) 
ggplot(subset(nullzoom, reps %in% levels(nullzoom$reps)[1:9])) + geom_line(aes(time, value)) + facet_wrap(~reps, scales="free")
````

Compute model-based warning signals on all each of these.  

``` {r summarize_null}
library(data.table)
library(reshape2)
library(earlywarning)
library(ggplot2)
#load("zoom.rda")
dt <- data.table(subset(nullzoom, value>250))
var <- dt[, warningtrend(data.frame(time=time, value=value), window_var), by=reps]$V1
acor <- dt[, warningtrend(data.frame(time=time, value=value), window_autocorr), by=reps]$V1
dat <- melt(data.frame(var=var, acor=acor))
ggplot(dat) + geom_density(aes(value), bw=0.2) + facet_wrap(~variable) + xlim(c(-1, 1))
````


