---
layout: post
category: ecology
tags:
- prosecutor

---


A quick foray into trying to understand why I see the wider distribution (though still symmetric) in the null of the OU model then in the null from the Allee model in the Prosecutor's fallacy.  

```{r, include=FALSE}
require(ggplot2)
require(reshape2)
require(data.table)
require(earlywarning)
```


```{r include=FALSE}
opts_knit$set(upload.fun = socialR::flickr.url)
opts_chunk$set(dev.args=list(bg="transparent"),
  tidy=FALSE, comment=NA, message=FALSE)
theme_notebook <- theme_grey() + 
  theme(plot.background = 
  element_rect(fill = "transparent",colour = NA),
  panel.grid.minor = 
  element_line(colour = "transparent"))
theme_set(theme_notebook)
```


Load the original run of the [ou model](https://github.com/cboettig/earlywarning/blob/712b63d2719839b6bbf4188d9affb34d7cec68ce/inst/examples/beer.md) and increase the `nulldt` data frame to use all points instead of a sample of length 5000

```{r}
load("beer_run.rda")

ou_dat <- dat
ou_null <- nulldat
#ou_null_ts <- nulldt

null <- timeseries #[1000:6010,]
null <- as.data.frame(cbind(time = 1:dim(null)[1], null))
ndf <- melt(null, id="time")
names(ndf) = c("time", "reps", "value")
levels(ndf$reps) <- 1:length(levels(ndf$reps)) # use numbers for reps instead of V1, V2, etc
nulldt <- data.table(ndf)
ou_null_ts <- nulldt
```

Plot the final distribution of indicator statistics, showing the width of the null,

```{r}
ggplot(dat) + geom_histogram(aes(value, y=..density..), binwidth=0.3, alpha=.5) +
 facet_wrap(~variable) + xlim(c(-1, 1)) + 
  geom_density(data=nulldat, aes(value), adjust=2) + xlab("Kendall's tau") + theme_bw()
```


Load the [allee model](https://github.com/cboettig/earlywarning/blob/712b63d2719839b6bbf4188d9affb34d7cec68ce/inst/examples/comment.md) and rename variables appropriately,

```{r}
load("comment_run.rda")
allee_dat <- dat
allee_null <- nulldat
allee_null_ts <- nulldt
```

For the Allee model, also plot the final distribution of indicator statistics, showing the width of the null,

```{r}
ggplot(dat) + geom_histogram(aes(value, y=..density..), binwidth=0.3, alpha=.5) +
 facet_wrap(~variable) + xlim(c(-1, 1)) + 
  geom_density(data=nulldat, aes(value), adjust=2) + xlab("Kendall's tau") + theme_bw()
```


Tidy the data and plot a single replicate from each.  Note the OU process has the correspondingly much wider null distribution due to 

```{r}
allee_x <- subset(allee_null_ts, reps==1)
ou_x <- subset(ou_null_ts, reps==1)
allee_x <- data.frame(time = allee_x$time, value = allee_x$value)
ou_x <- data.frame(time = ou_x$time, value = ou_x$value)

ggplot(allee_x, aes(time, value)) + geom_point()
ggplot(ou_x, aes(time, value)) + geom_point()

```



Note that subsampling the data at coarser interval doesn't matter

```{r}
warningtrend(ou_x[seq(1, length(ou_x$value),by=50),], window_var)
```

Nor does scaling matter (recall Kendall's $tau$ is a rank-correlation test)

```{r}
warningtrend(data.frame(time=ou_x$time, value=(ou_x$value+1)*500), window_var)
warningtrend(data.frame(time=ou_x$time, value=ou_x$value), window_var)
```


Lengthening the sample suggests the sampling is not long enough to have converged in distribution of this statistic.  Computing on this fine sampling resolution over adequate length of time quickly becomes prohibitive.   

```{r}
sapply(seq(1000, 20000, by=1000), function(i) warningtrend(ou_x[1:i,], window_var))

```
